Trans-Dimensional Bayesian Inference for Gravitational Lens Substructures

### 2015-09-09 Papers/ms.tex
Include the mu_lens figure
new numbers

@@

We generated 5,000 samples from the DNS target distribution and thinned
by a factor of $10^5$, so $5 \times 10^8$ MCMC iterations were actually
performed, taking approximately 40 hours. The first 1/3 of the run was
excluded as burn-in. After resampling, this resulted in 632
equally weighted posterior samples. Although this image is smaller than the
simulated data ($46 \times 46$ pixels), we fired $2 \times 2$ rays per pixel
for greater accuracy.

@@

The estimated marginal likelihood of our model
is
$\ln\left[p(\boldsymbol{D} | \boldsymbol{\theta})\right] \approx -6967$, and
the information (Kullback-Leibler divergence from the prior to the posterior)
is $\mathcal{H} \approx 129$ nats. 
Three example models (lens and source) sampled from the posterior are shown
in Figure~\ref{fig:sources}. {\revisions
The model was able to fit the data down to the noise level.
}

@@

\subsection{Reproducibility of the results}\label{sec:convergence}
The ``effective sample size'' returned by DNS, which we have described as the
number of posterior samples, takes into account the fact DNS's target
distribution is not the posterior. However, it does not take autocorrelation
into account, and thus can present an optimistic picture of the accuracy of
any Monte Carlo approximations to posterior quantities. Most standard diagnostic
techniques used in MCMC can be applied here. The simplest of these is a
check of reproducibility. If different runs yield substantially different
results, the MCMC output should be treated with caution. For example, in
Figure~\ref{fig:masses3}, there is a correlation between the mass attributed
to the SIE component and that attributed to substructures. However, the
distribution also appears somewhat ``lumpy'' or multimodal. This may
not be a feature of the actual posterior distribution, but could arise due to
imperfect sampling. Whereas a standard Metropolis sampler would move around
very slowly in the hypothesis space, DNS naturally spends a non-negligible fraction
of the time sampling the prior. Therefore, a particle exploring the parameter
space can ``forget'' its good fitting position, move somewhere completely
different, and find another good fitting model in a different location. This
is a natural feature of the algorithm (and is also present in related
algorithms such as parallel tempering). Despite this, the patchy
nature of Figure~\ref{fig:masses3} suggests that posterior exploration is
still challenging in this problem.

>>>>

\begin{figure}
\centering
\includegraphics[scale=0.45]{mu_lens.pdf}
\caption{{\revisions The posterior distribution for $\mu_{\rm lens}$, the
hyperparameter which determines the prior expected value of substructure
masses, given the `harder' simulated data. The true value (in arbitrary units)
was 1, which is indicated by the vertical line.
}
\label{fig:mu_lens}}
\end{figure}

@@

We generated 5,000 samples from the DNS target distribution and thinned
by a factor of $10^5$, so $5 \times 10^8$ MCMC iterations were actually
performed, taking approximately 40 hours. After resampling, this resulted in
{\revisions 718}
equally weighted posterior samples. Although this image is smaller than the
simulated data ($46 \times 46$ pixels), we fired $2 \times 2$ rays per pixel
for greater accuracy.

@@

The estimated marginal likelihood of our model
is {\revisions
$\ln\left[p(\boldsymbol{D} | \boldsymbol{\theta})\right] \approx -6948.5$}, and
the information (Kullback-Leibler divergence from the prior to the posterior)
is {\revisions $\mathcal{H} \approx 123.1$} nats. 
Three example models (lens and source) sampled from the posterior are shown
in Figure~\ref{fig:sources}. {\revisions
The model was able to fit the data down to the noise level.
}

@@

\subsection{Reproducibility of the results}\label{sec:convergence}
The ``effective sample size'' returned by DNS, which we have described as the
number of posterior samples, takes into account the fact DNS's target
distribution is not the posterior. However, it does not take autocorrelation
into account, and thus can present an optimistic picture of the accuracy of
any Monte Carlo approximations to posterior quantities. Most standard diagnostic
techniques used in MCMC can be applied here. The simplest of these is a
check of reproducibility. If different runs yield substantially different
results, the MCMC output should be treated with caution. For example, in
Figure~\ref{fig:masses3}, there is a correlation between the mass attributed
to the SIE component and that attributed to substructures. However, the
distribution also appears somewhat ``lumpy'' or multimodal. This may
not be a feature of the actual posterior distribution, but could arise due to
imperfect sampling. Whereas a standard Metropolis sampler would move around
very slowly in the hypothesis space, DNS naturally spends a non-negligible fraction
of the time sampling the prior. Therefore, a particle exploring the parameter
space can ``forget'' its good-fitting position, move somewhere completely
different, and find another good-fitting model in a different location. This
is a natural feature of the algorithm (and is also present in related
algorithms such as parallel tempering). Despite this, the patchy
nature of Figure~\ref{fig:masses3} suggests that posterior exploration is
still challenging in this problem.
